---
title: "Distribution of the power in the CDSR"
author: "Erik van Zwet"
header-includes:
  - \usepackage{amsmath}
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  pdf_document:
    toc: true
    toc_depth: 3
    number_sections: true
---

<!--- code from Erik for creating power curve --->

# Get the data

Get the data from Schwab, S. (2020) Re-estimating 400,000 treatment effects from intervention studies in the Cochrane Database of Systematic Reviews [Data set]. Open Science Framework. (https://osf.io/xjv9g/).


```{r, warning=FALSE, message=FALSE, echo=TRUE}
library(ggplot2)
library(dplyr)
library(flexmix)  # for estimating normal mixture distribution

d=read.csv("https://osf.io/xq4b2/?action=download")
d=d %>% filter(RCT=="yes")
d=d[d$outcome.group=="efficacy" & d$outcome.nr==1 & abs(d$z)<20,]
d=group_by(d,study.id.sha1) %>% sample_n(size=1)   # select single outcome per study 

z=d$z
cat("number of z-values:",length(z),"\n")
cat("proportion significant:",mean(abs(z)>1.96),"\n")
```

# Mixture distribution of z and SNR

Estimate the distribution of the z-values with a 4-part normal mixture. We restrict the means to be zero, as we also did in van Zwet, E., Schwab, S. and Senn, S., 2021. The statistical properties of RCTs and a proposal for shrinkage. _Statistics in Medicine_, 40(27), pp.6107-6117.

```{r, warning=FALSE, message=FALSE, echo=TRUE}
fit=flexmix(z ~ 0, k = 4)        # estimate distr of b/s
for (i in 1:5){                  # try few restarts
  fit1=flexmix(z ~ 0, k = 4)
  if (fit1@logLik > fit@logLik) fit = fit1
}

p=summary(fit)@comptab$prior
s=parameters(fit)
```

Do the "deconvolution trick" to get the standard deviations of the SNR.

```{r, warning=FALSE, message=FALSE, echo=TRUE}
s=sqrt(s^2 - 1)              # standard deviations of SNR
```

#  Distribution of the power

Sample 10^6 from the mixture distribution of the SNR, and transform them to the power.

```{r, warning=FALSE, message=FALSE, echo=TRUE}
rmix = function(n,p,m,s){   # sample from a normal mixture
  d=rmultinom(n,1,p)
  rnorm(n,m%*%d,s%*%d)
}

snr=rmix(10^6,p,m=rep(0,4),s)
power=pnorm(-1.96,snr,1) + 1 - pnorm(1.96,snr,1)
summary(power)
```

Finally, make a histogram of the power.

```{r, warning=FALSE, message=FALSE, echo=TRUE}
df=data.frame(power=power)
ggplot(df, aes(x=power, y=..density..)) +
  geom_histogram(bins=40,fill="white",col="black") + 
  scale_x_continuous(breaks = seq(0,1,0.1)) +
  theme_bw() + labs(x="power",y="")

# added by AGB
median = median(df$power)*100
text = data.frame(x=median, y=200000, label=paste('Median = ', round(median), '%', sep=''))
gplot = ggplot(df, aes(x=power*100)) +
  geom_vline(xintercept = median, lty=2)+
  geom_label(data=text, aes(x=x, y=y, label=label), adj=-0.1)+
  geom_histogram(bins=40, fill="skyblue", col="grey33") + 
 # scale_x_continuous(breaks = seq(0,1,0.1)*100) +
  theme_bw() + 
  theme(panel.grid.minor = element_blank())+
  labs(x="Power (%)",y="Count")
ggsave(filename = 'figures/power_histogram.jpg', width=5, height=4, units='in', dpi=500)
```
